---
title: "PA1_template"
author: "J.Roland"
date: "5/5/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Loading and preprocessing the data
Read in the data
```{r}
dataset <- read.csv("activity.csv")
```

Change date string to date format  
```{r}
dataset$date <-(as.Date(dataset$date))
```

Get ggplot2 library
```{r}
library(ggplot2)
```
## What is mean total number of steps taken per day?

Get sum of steps per day
```{r}
dailySteps <- tapply(dataset$steps, dataset$date, FUN=sum, na.rm=TRUE)
```

Make a histogram of steps per day
```{r}
hist(dailySteps,col="blue")

```

Calculate the mean steps per day
```{r}
dailyStepsMean <-mean(dailySteps, na.rm=TRUE)
dailyStepsMean
```

Calculate the median steps per day
```{r}
dailyStepsMedian <-median(dailySteps, na.rm=TRUE)
dailyStepsMedian
```
##What is the average daily activity pattern?

calculate mean of steps per interval
```{r}
average_steps <- aggregate(x=list(steps=dataset$steps), by=list(interval=dataset$interval),FUN=mean, na.rm=TRUE)
```

plot it
```{r}
ggplot(data=average_steps, aes(x=interval, y=steps, colour="#000099")) + geom_line() + ggtitle("Time Series: average number of steps") + xlab("5-minute interval") + ylab("average number of steps taken")
```

identify interval with the highest average max # of steps
```{r}
average_steps[which.max(average_steps$steps),]
```

##Imputing missing values

count the number of NAs (missing data)
```{r}
sum(is.na(dataset$steps))
```

replace NA values with the mean of the steps
```{r}
dataset$steps[is.na(dataset$steps)]<-mean(dataset$steps,na.rm=TRUE)
```

Get sum of steps per day with the NA values replaced
```{r}
replacedNA_dailySteps <- tapply(dataset$steps, dataset$date, FUN=sum, na.rm=TRUE)
```
Make a histogram of steps per day with the NA values replaced
```{r}
hist(replacedNA_dailySteps,col="blue")
```

Calculate the new "replacedNA" mean steps per day
```{r}
replacedNA_dailyStepsMean <-mean(replacedNA_dailySteps, na.rm=TRUE)

replacedNA_dailyStepsMean
```
Calculate the new "replacedNA" median steps per day
```{r}
replacedNA_dailyStepsMedian <-median(replacedNA_dailySteps, na.rm=TRUE)

replacedNA_dailyStepsMedian
```

##Are there differences in activity patterns between weekdays and weekends?

change date format to weekdays
```{r}
dataset$days=weekdays(dataset$date)
```

change day names to weekend or weekday
```{r}
dataset$days_type<-ifelse(dataset$days=="Saturday"|dataset$days=="Sunday","weekend","weekday")
```

calculate mean steps taken per interval, using the weekday/weekend categorization
```{r}
average_steps_byDaysType<-aggregate(dataset$steps,by=list(dataset$interval,dataset$days_type),FUN=mean,na.rm=TRUE)
```

simplify the column names in preparation for plotting
```{r}
colnames(average_steps_byDaysType)<-c("interval","day_type","steps")
```

plot the graphs of weekend versus weekday steps by interval
```{r}
ggplot(aes(x=interval,y=steps),data=average_steps_byDaysType)+geom_line()+facet_wrap(~average_steps_byDaysType$day_type)
```












